Title

### 2017-04-09 paper.tex
Preliminary descriptions

@@

    Identifying that scheduler is also a source of jitter, we applied our proposed scheduler profiling tool to evaluate
    and identify of which part of the scheduling process is the source of jitter. Knowing the IRQ latency, and
    context-switching latency, we would like to evaluate the scheduler duration. Traditionally, the latency of this part
    is considered static with the CFS algorithm complexity, designed to be O(1), but real measurement result are
    lacking. Thus, we set our proposed profiler to record the increment of scheduler run-queue and scheduler
    context-switching events with switched-to PID of each processors. A analyzer will process the recorded data and
    gives out the scheduler duration of a specified process, enabling us the understand the scheduling of real-time task
    in SMP system. The result of running a solo instance of cyclictest on SoC FPGA Development Kit is shown in figure
    \ref{fig:sd_solo}. The latency could be considered static with maximum 17us. This result matches traditional
    considerations, but with the interference from other work-load, the jitter will start to emerge.

    \begin{figure} \centering \includegraphics[width=3.5in]{img/sd-solo.png} \caption{Scheduler duration of solo
    cyclictest, no load} \label{fig:sd_solo} \end{figure}

\subsubsection{Overloaded System}

    We created additional work-load, executing alone with cyclictest during profiling session. We created a mild loading
    condition, which has 3 stress threads, and a heavy loading condition, which has 8 stress threads. The result of mild
    loading condition is shown in figure \label{fig:sd_mild}. This result also exhibits a mostly static scheduler
    duration with seldom jitter with maximum of 20us. But putting the system in to heavy loading condition, the result,
    shown in figure \ref{fig:sd_heavy} exhibits random jitter up to 34us. The spread of distribution of scheduler
    duration, shown in figure \ref{fig:sd_heavy_d}, would become wilder. This informs us that the scheduler duration is
    not a static value and will increase alone with the system's load. Also informs that the scheduler duration contains
    but not only contains CFS algorithm task selection process which is proved to be O(1).

    \begin{figure} \centering \includegraphics[width=3.5in]{img/sd-mild.png} \caption{Scheduler duration of cyclictest
    with mild load} \label{fig:sd_mild} \end{figure}

    \begin{figure} \centering \includegraphics[width=3.5in]{img/sd-heavy.png} \caption{Scheduler duration of cyclictest
    with heavy load} \label{fig:sd_heavy} \end{figure}

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-heavy-d.png} \caption{Scheduler duration distribution
    of cyclictest with heavy load} \label{fig:sd_heavy_d} \end{figure}

\subsubsection{Short Inter-Arrival Time}

    We also captured large jitter of scheduler duration on a mild loaded system. The experiment is running mctest with
    cyclictest. We pinned kernel module mctest to use first processor, causing cyclictest real-time thread and main
    thread being scheduler onto second processor. On most of the time, this execution schema results the same result as
    cyclictest executing solely in the system. But on some occasions, the cyclictest main thread would be waken-up and
    enter run-queue with short inter-arrival time against the cyclictest real-time thread, which was also being
    waken-up. When this event happened, the scheduler duration will raise to maximum of 23us. This result, shown in
    figure \ref{fig:sd_mctest} hints us that not only long term overload system would exhibit large scheduler duration,
    but also short burst a overloading condition during the real-time task scheduling and execution would cause the
    scheduler duration to raise.

    \begin{figure} \centering \includegraphics[width=3.5in]{img/sd-mctest.png} \caption{Scheduler duration of cyclictest
    with short inter-arrival time to run-queue against other task} \label{fig:sd_mctest} \end{figure}

>>>>

The processing of time-critical tasks, depends upon the capacity of the system to respondto an event, within
a known and bounded interval of time.

Existing techniques for effectively scheduling real-time workloads on multiprocessors has been increasing recently.
Two factors are driving this movement. First, algorithmic research on this topic has resulted in a number of new
scheduling approaches that remove some of the fundamental barriers imposed by partitioning approaches, the scheme
smost commonly considered in earlier work. Second, the landscape in terms of hardware platforms has been changing:
"server-class" multiprocessor machines have been availablefor some time now, and SoC vendors are shifting to
multi-core technologies as a solution to the "thermal roadblock" imposed by single-core designs. In multicore
platforms, several processor cores are placed on a single chip. Most major chip manufacturers currently offer
more core.

@@

    Identifying that scheduler is also a source of jitter, we applied our proposed scheduler profiling tool to evaluate
    and identify of which part of the scheduling process is the source of jitter. Knowing the IRQ latency, and
    context-switching latency, we would like to evaluate the scheduler duration. Traditionally, the latency of this part
    is considered static with the CFS algorithm complexity, designed to be O(1), but real measurement result are
    lacking. Thus, we set our proposed profiler to record the increment of scheduler run-queue and scheduler
    context-switching events with switched-to PID of each processors. A analyzer will process the recorded data and
    gives out the scheduler duration of a specified process, enabling us the understand the scheduling of real-time task
    in SMP system. The result of running a solo instance of cyclictest on SoC FPGA Development Kit is shown in figure
    \ref{fig:sd_noload}. The latency could be considered static with maximum 17us. This result matches traditional
    considerations, but with the interference from other work-load, the jitter will start to emerge.

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-noload.png} \caption{Scheduler duration of solo
    cyclictest, no load} \label{fig:sd_noload} \end{figure}

\subsubsection{Scheduler Duration under Different Loading Condition}

    To measure the source of scheduler duration jitter, We created additional work-load, executing alone with
    cyclictest, schedulered with schedFIFO with highest priority in SoCFPGA based system, during profiling session. We
    created a overloaded condition, which has 8 stress threads, doing either CPU, memory or I/O bounded task. The
    scheduler duration of cyclictest real-time task under each type of loading condition is shown in figure
    \ref{fig:sd_cpu}, \ref{fig:sd_memory}, and \ref{fig:sd_io}. This result exhibits a mostly static scheduler
    duration with seldom jitter with maximum of 20us in CPU and I/O bounded loads. But putting the system in to heavy
    memory loading condition, the result, shown in figure \ref{fig:sd_memory} shows an increased average of scheduler
    duration and exhibits random jitter up to 34us. The spread of distribution of scheduler duration would also become
    wilder. This informs us that the scheduler duration is not a static value and will increase alone with the system's
    load, especially memory related load. This informs that the scheduler runnable task selection process execution time
    could be lengthen by user-space task, doing memory bounded tasks.

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-cpu.png} \caption{Scheduler duration of cyclictest
    with CPU bounded load} \label{fig:sd_cpu} \end{figure}

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-memory.png} \caption{Scheduler duration of cyclictest
    with memory bounded load} \label{fig:sd_memory} \end{figure}

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-io.png} \caption{Scheduler duration distribution of
    cyclictest with I/O bounded load} \label{fig:sd_io} \end{figure}

    To mitigate the influence of memory bounded tasks, we isolated the real-time task into dedicated CPU core, and
    pinned memory bounded tasks to another CPU core. In addition, we make the shared L2 cache only accessible from the
    CPU core, running real-time task. The result of scheduler duration in shown in figure \ref{fig:sd_iso}. The
    scheduler duration is stabilized and showing as distribution as other loading conditions. The stress memory bounded
    tasks were designed to spin on malloc and free calls. This suggests us that memory allocation and deallocation would
    decrease the scheduler performance and increase its execution jitter via cache and memory model. Also, we gathered
    the result of pining both real-time and loads to same CPU core. Figure \ref{fig:sd_same_no_cache} shows the result
    of running on CPU core without L2 cache.  Comparing with figure \ref{fig:sd_memory}, result under 2 core SMP, this
    result showns a slight improvement of average latency and jitter. This informs that shared L2 cache and dynamic
    number of loading tasks, sharing CPU with real-time task, would be the source of scheduler duration jitter. Thus,
    decrease the use of dynamic memory allocation, real-time task CPU isolation and cache lockdown could prevent raising
    the scheduler duration, keeping it as a stable value, about 10us.

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-iso.png} \caption{Scheduler duration of cyclictest with
    memory bounded load on different cores} \label{fig:sd_iso} \end{figure}

    \begin{figure} \centering \includegraphics[width=3in]{img/sd-same-no-cache.png} \caption{Scheduler duration of
    cyclictest with memory bounded load on same core} \label{fig:sd_same_no_cache} \end{figure}
